---
title: "conservation areas exploration"
author: "Iris Foxfoot"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(here)
library(tidyverse)
library(janitor)
library(raster)
library(fasterize)
```

This is an exploration of the California Conservation Easement Database (CCED) and California Protected Area Database (CPAD). The goal is to find conservation areas that have never been farmed, which are similar enough to compare to agricultural areas

## Read in All Data

```{r, read in data}
#read in cced database
easements <- read_sf(here("data", "CCED_2021b", "CCED_2021b_Release.shp"))

#read in cpad database
protected_areas_holdings <- read_sf(here("data", "CPAD_2021b", "CPAD_2021b_Holdings.shp"))
protected_areas_superunits <- read_sf(here("data", "CPAD_2021b", "CPAD_2021b_SuperUnits.shp"))
protected_areas_units <- read_sf(here("data", "CPAD_2021b", "CPAD_2021b_Units.shp"))

#getting crs for protected ares
proj_crs <- st_crs(easements)

#read in aggregated agriculture file (years 2016-2019), transform to same crs as other files
farms <- read_sf(here("data", "shapefiles_written", "unioned_fields_20162019", "unioned_fields_20162019.shp")) %>% 
  st_transform(crs = proj_crs)

#read in kern county shapefile
kern_county <- read_sf(here("data", "Kern_County_Boundary")) %>% 
  st_transform(crs = proj_crs)

#read in soil data
soil_quality <- read_sf(here("R_input", "spatial", "SSURGOsoil", "kern_soil.gpkg")) %>% 
  st_transform(crs = proj_crs)

soil_quality_factX <- read_sf(here("R_input", "spatial", "SSURGOsoil", "kern_soil_factX.gpkg")) %>% 
  st_transform(crs = proj_crs)

#read in historical landuse data
hist_landu <- read_sf(here("data", "esrp_histlu", "histlu200408.shp")) %>% 
  st_transform(crs = proj_crs) %>% 
  dplyr::select(POT_LU, LU1885, LU1912, LU1945, LU1977, LU2004)
```

## Narrow Down Conservation Areas

Select only conservation areas within Kern county. Then select conservation areas with a GAP 1, 2, or 3 status OR areas that were established before 1950. 

```{r}
#get protected areas within kern county
easements_kern <- easements %>% 
  filter(st_intersects(geometry, kern_county, sparse = FALSE))

holdings_kern <- protected_areas_holdings %>% 
  filter(st_intersects(geometry, kern_county, sparse = FALSE))
  
superunits_kern <- protected_areas_superunits %>% 
  filter(st_intersects(geometry, kern_county, sparse = FALSE))

units_kern <- protected_areas_units %>% 
  filter(st_intersects(geometry, kern_county, sparse = FALSE))

#get gap status 1, 2, 3, or pre 1950s conservation areas
#also make col names consistent

long_easements <- easements_kern %>% 
  filter(gapcat <= 3 | year_est %in% c(1:1950)) %>% 
  clean_names() %>% 
  mutate("unit_id" = e_hold_id,
         "unit_name" =  sitename,
         "agncy_name" = esmthldr,
         "yr_est" =  year_est,
         type = "easement") %>% 
  dplyr::select(type, unit_id, unit_name, agncy_name, yr_est, gap1_acres, gap2_acres, gap3_acres, gap4_acres)

long_holdings <- holdings_kern %>% 
  filter(GAP1_acres > 0 | GAP2_acres > 0 | GAP3_acres > 0 | YR_EST %in% c(1:1950)) %>% 
  clean_names() %>%
  mutate(type = "holding") %>% 
  dplyr::select(type, unit_id, unit_name, agncy_name, yr_est, gap1_acres, gap2_acres, gap3_acres, gap4_acres)

long_superunits <- superunits_kern %>% 
  filter(GAP1_acres > 0 | GAP2_acres > 0 | GAP3_acres > 0 | YR_EST %in% c(1:1950)) %>% 
  clean_names() %>% 
  mutate("unit_id" = suid_nma,
         "unit_name" =  park_name,
         "agncy_name" = mng_agency,
         type = "superunit") %>% 
  dplyr::select(type, unit_id, unit_name, agncy_name, yr_est, gap1_acres, gap2_acres, gap3_acres, gap4_acres)
  
long_units <- units_kern %>% 
  filter(GAP1_acres > 0 | GAP2_acres > 0 | GAP3_acres > 0 | YR_EST %in% c(1:1950))  %>% 
  clean_names() %>%
  mutate(type = "unit") %>% 
  dplyr::select(type, unit_id, unit_name, agncy_name, yr_est, gap1_acres, gap2_acres, gap3_acres, gap4_acres)

#bind together 
#not including units/superunits because they are super big aggregations
kern_conservation_areas <- rbind(long_easements, long_holdings) %>% 
  mutate(id = rownames(.))
```

## Add Historical Landuse Data

Take the resulting conservation areas and merge it with historical landuse data, so you can see if the conservation area was previously irrigated. 

```{r}
#explore historical landuse in the conservation area
kern_conserve_hist <- st_join(kern_conservation_areas, hist_landu, left = T)
```

## Add Soil Data

```{r}
#explore soil quality of conservation areas by joining the two shapefiles
kern_conserve_soil_joined <- st_join(kern_conserve_hist, soil_quality, left = T)
```

## Filter

```{r}
#filter for areas with good soil, in valley, and not irrigated in 2004
ideal_conservation_area <- kern_conserve_soil_joined %>% 
  filter(interphrc %in% c("Grade 1 - Excellent", "Grade 2 - Good")) %>% 
  filter(!is.na(POT_LU)) %>% 
  # filter(LU2004 != "Irrigated") %>% 
  # filter(LU1977 != "Irrigated")

#get polygons within three km of 2016-2019 farms
ideal_conservation_near_farms <- st_intersection(ideal_conservation_area, st_buffer(farms, 3000))

#add column for historical irrigation

#taking a loooong time
#ideal_conservation_not_farmed <- st_difference(ideal_conservation_near_farms, farms)

#write to shapefile
#st_write(ideal_conservation_not_farmed, "data/shapefiles_written/conservation_areas_may18.shp")

#get original data that matches the id of polygons in ideal conservation near farms
best_conserve_bets <- kern_conservation_areas %>% 
  filter(id %in% ideal_conservation_near_farms$id)

#group to get list of names.
best_conserve_bets_grouped <- best_conserve_bets %>% 
  group_by(unit_id, unit_name) %>% 
  count()
```

