---
title: "Kern County Fields Through Time"
author: "Iris Foxfoot"
date: "1/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r get_packages}
knitr::opts_chunk$set(echo = TRUE)
library(sf) #for polygons
library(tidyverse) #for data wrangling
library(janitor) #for converting names to lowercase snake
library(tmap) #for interactive maps
library(reshape2) #for melting and recasting data frames (long to wide format)
```

## Read in field data and bind it together

In this section read in the field data and wrangle it such that the end product is a single SF object where each row is a field and the columns are for geometry and for permitsite ID, crop code, and year, which are merged to single string.

```{r field_data_wrangling}
#read in kern data
#must have the following file path in working directory: data/kernfields/kernYEAR.shp

#create list of years
kern_field_years <- c(1997:2021)

#loop through read-in process
for (i in kern_field_years) {
  filename <- paste0("kernfields_", i)
  wd <- paste0("data/kernfields/kern", i, ".shp")
  assign(filename, read_sf(wd))
}

kernfields_2021 <- kernfields_2021 %>% 
  mutate(comm = Commodity) %>% 
  mutate(comm_code = COMM_COD_1)

#loop through listed sf objects and tidy
fields_cleaned_list <- lapply(mget(sprintf("kernfields_%d", 1997:2021)), 
                 function(x) 
                   {
                   clean_names(x) %>%
                     st_make_valid() %>% 
                     st_transform(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>%
                     #select pmt_site and comm (not all years have comm code)
                     dplyr::select(pmt_site, comm)
                     }
                 )


#bind crop data together, include new column "source" saying what dataset it came from (year)
kernfield_bind <- do.call(rbind, lapply(names(fields_cleaned_list), 
                                    function(x) 
                                      cbind(fields_cleaned_list[[x]], 
                                            source = x)
                                    ))
```

## Fix crop codes so they are consistent accross years 

```{r field_data_wrangling}
#fixing commodity codes (only in post-2017 data)

#get unique commodities from all years
unique_crops <- as.data.frame(unique(kernfield_bind$comm)) %>% 
  rename(comm = `unique(kernfield_bind$comm)`)

#get comm codes and crops from 2017-2019
comm_codes_pre_20 <- rbind(kernfields_2017, kernfields_2018, kernfields_2019, kernfields_2019) %>% 
  clean_names() %>% 
  st_drop_geometry() %>%
  select(comm_code, comm)

#comm codes from 2020
comm_2020 <- kernfields_2020 %>% 
  select(COMM_CODE, COMM) %>%
  st_drop_geometry() %>%
  clean_names()

#comm codes from 2021
comm_2021 <- kernfields_2021 %>% 
  select(comm_code, comm) %>%
  st_drop_geometry() %>%
  clean_names()

#bind them together, find every unique comm and its matching code
comm_codes_all <- rbind(comm_codes_pre_20, comm_2020, comm_2021) %>% 
  group_by(comm) %>% 
  unique()

#left join codes with crops
comm_with_crops <- left_join(unique_crops, comm_codes_all) %>% 
  mutate(comm_code = case_when(comm == "ALFALFA SEED" ~ "23001", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "ARRUGULA" ~ "13056", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "BARLEY FOR/FOD-" ~ "22008", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "BARLEY SEED" ~ "22008", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "BEAN DRIED SEED" ~ "15001", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "BEAN SUC SEED" ~ "15003", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "BEET SEED" ~ "29109", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "CAULIFLOWR SEED" ~ "13008", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "CELERY SEED" ~ "29113", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "CORN SEED" ~ "24002", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "GRAPE (RAISIN)" ~ "29141", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "GRAPE (WINE)" ~ "29143", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "MELON SEED" ~ "29122", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "OAT SEED" ~ "29125", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "ONION GR SEED" ~ "16004", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "PARSLEY SEED" ~ "13022", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "RADISH SEED" ~ "14014", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "SAFFLOWER SEED" ~ "29129", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "SORGHUM SEED" ~ "29131", T ~ comm_code)) %>%
  mutate(comm_code = case_when(comm == "WHEAT SEED" ~ "29139", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "COTTON FOR/FOD" ~ "29121", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(comm == "OF-DEC. TREE" ~ "152", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(str_detect(comm, "OP-") ~ "154", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(str_detect(comm, "OT-") ~ "156", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(str_detect(comm, "PEAS ") ~ "29127", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(str_detect(comm, "TURNIP") ~ "29137", T ~ comm_code)) %>% 
  mutate(comm_code = case_when(str_detect(comm, "VEGETABLE") ~ "13000", T ~ comm_code))

#replace all remaining NA codes with number starting with 99
comm_with_crops$comm_code[is.na(comm_with_crops$comm_code)] <- sample(9900:9999, size = sum(is.na(comm_with_crops$comm_code)), replace = F)

#make sure chinese green codes are same for seed and conventional version
comm_with_crops <- comm_with_crops %>% 
  mutate(comm_code = case_when(comm == "CHINESE GR SEED" ~ "9998", T ~ comm_code))

#merge codes to field data, now each observation has a crop code
kernfield_comm_code <- merge(kernfield_bind, comm_with_crops, by = "comm")

#now create column with year, pmt_site, id, and crop code
kernfield_clean <- kernfield_comm_code %>%
  #turn source column into year column by keeping only numbers
  mutate(year = gsub("\\D", "", source)) %>% 
  #Create ID column
  mutate(id = row_number()) %>% 
  #merge comm code to pmt_site
  mutate(pmt_site_code = str_c(pmt_site, "-", comm_code)) %>% 
  #merge ID to  pmt_site
  mutate(pmt_site_id = str_c(pmt_site_code, "-", id)) %>% 
  #merge this column with year
  mutate(pmt_site_string = str_c(year, "_", pmt_site_id)) %>% 
  #select only column we need, clean geometry
  dplyr::select(pmt_site_string) %>% 
  st_buffer(dist = 0) %>% #for some reason it only works with a zero buffer
  st_make_valid()

#write to shapefile
#st_write(kernfield_clean, "data/shapefiles_written/1997-2021kernfields.shp", driver = "ESRI Shapefile")
```

Take the resulting shapefile `sample_clean.shp` and upload it to mapshaper's online interface (https://mapshaper.org/). Select the box saying "detect line intersections" and the box "snap vertices". Then enter in the console `$ mosaic`. The processing can take 30-45 minutes. Download results as a shapefile by clicking export and selecting shapefile. It will download in a zipfile.

This thread could potentially improve workflow https://github.com/mbloch/mapshaper/issues/353 

## Create pivot table using mapshaper 
(where col name is from year and rows are smallest intersection of fields)

```{r}
#read in data from mapshaper outputs
mosaic <- read_sf("data/mapshaper_outputs/1997-2021kernfields/1997-2021kernfields.shp") 

#drop polys under one acre and clean
mosaic_drop_acre <- mosaic %>%
  mutate(area = st_area(.)) %>% 
  filter(area >= units::set_units(43560,"ft^2")) %>% 
  st_transform(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  st_make_valid()

#negative buffer and intersect with field data
kernfield_join <- st_intersection(st_buffer(mosaic_drop_acre, dist = -1), kernfield_clean)

#pivot so each column is a year
kernfield_pivot <- kernfield_join %>%
  filter(!is.na(pmt_site_string)) %>% 
  separate(pmt_site_string, into = c("year", "pmt_site"), sep = "_") %>% 
  select(-c(FID, area)) %>% 
  pivot_wider(names_from = year, values_from = pmt_site) %>%
  st_sf() %>% 
  st_buffer(dist = 1)

#drop shapes under an acre again
kernfield_pivot_over_acre <- kernfield_pivot %>%
  mutate(area_meters_sq = st_area(.)) %>% 
  filter(area_meters_sq >= units::set_units(1,"acre"))

#converting list to strings of characters so I can save to csv
#kernfield_pivot_over_acre = data.frame(lapply(kernfield_pivot_over_acre, as.character), stringsAsFactors=FALSE)

#write to csv
#write_csv(kernfield_pivot_over_acre, "data/shapefiles_written/1997-2021_pivoted.csv")
```

## Calculate fallowed stats

In this section I calculate the total number of years fallowed, the greatest number of years fallowed in a row, the start of the streak, the end of the streak, and the last active crop before the streak

```{r}
#read in csv
#pivot_table <- read_csv("data/shapefiles_written/1997-2021_pivoted.csv")

#make sure cols are in order (from earliest year to latest year)
pivot_table <- kernfield_pivot_over_acre %>% 
  select(sort(names(.))) %>%
  #add id column
  mutate(id = rownames(.))

#make names numeric
#colnames(pivot_table) <- c("area_meters_sq", "geometry", 1997:2021, "id")

#create list of years
#kern_field_years <- c(1997:2021)

#create dataset for for for loop to loop over
fallowed_sample_pivot <- pivot_table

#for loop that creates a new column per each year, that indicated if field was fallowed or not
for (i in kern_field_years) {
fallowed_sample_pivot <- fallowed_sample_pivot %>%
  mutate(!!paste0("fallowed_", i) 
         := case_when(!!as.name(i) == "NULL" ~ "fallowed",
                      str_detect(!!as.name(i), "-66000-") & 
                        str_detect(!!as.name(i), ", ", negate = T) ~ "fallowed",
                            T ~ "active"))
}

#drop geometry
fallowed_sample_no_geo <- fallowed_sample_pivot %>% 
  st_drop_geometry()

#transform wide data to the long format
sample_long = melt(fallowed_sample_no_geo, id.vars = c("1997":"2021", "area_meters_sq", "id")) 

#write new column names
colnames(sample_long) = c("1997":"2021", "area_meters_sq", "id", "year", "status")

#calculate how long each field has been fallowed by counting repeated fallowed status within rows
sample_long_mutate = sample_long %>%
  #created p-columns that correspond to year
  mutate(p_var = paste0("years_fallowed_", substr(year, 10, 14))) %>% 
  group_by(id) %>%
  #create column that counts how many active or fallowed are in a row
  mutate(lapsed_num = sequence(rle(status)[["lengths"]]),
         #only keep fallowed streaks, if active it has been fallowed zero years so replace with zero
         final = ifelse(status == "active", 0, 
                           ifelse(str_detect(status, "fallowed"), 
                                  lapsed_num, NA))) %>% 
  select(-lapsed_num) %>% 
  data_frame()

#convert if from long format to wide format 
#NOTE: col names need to be changed when more years are added
df_p = dcast(sample_long_mutate[, c(1:29, 30:31)], id  ~ p_var,  value.var = "final")

#merge it with original data
years_fallowed = merge(df_p, fallowed_sample_pivot, id = id) %>%
  select(-c(starts_with("fallowed_"))) %>% 
  mutate(total_yrs_fallowed = rowSums(select(.,starts_with("years_fallowed"))!=0))

#create df of only fallow numbers
longest_fallow <- years_fallowed %>%
  select(starts_with("years_fallowed"))

#create a column indicating where the fallow streak ends (based on max value)
longest_fallow$end_fallow_streak<-colnames(longest_fallow)[apply(longest_fallow, 1, which.max)]

#calculate how long the fallowed streak was
longest_fallow_years <- longest_fallow %>%
  mutate(longest_fallow_streak = do.call(
    pmax, subset(., select = years_fallowed_1997:years_fallowed_2021))) %>% 
  mutate(end_fallow_streak = str_extract(end_fallow_streak, "\\d+")) %>%
  select(longest_fallow_streak, end_fallow_streak)

#add column that says start year of fallow crop
longest_fallow_years$start_fallow_streak <- 
  as.numeric(longest_fallow_years$end_fallow_streak) - 
  as.numeric(longest_fallow_years$longest_fallow_streak)
  
#bind these new columns with orginal data
years_fallowed_clean <- cbind(years_fallowed, longest_fallow_years) %>% 
  select(-starts_with("years_fallowed"))

#replaces year with last active crop
last_crop <- years_fallowed_clean %>%
  mutate(last_crop_before_streak = 
           case_when(start_fallow_streak == 1996 ~ "starts_fallowed",
                     start_fallow_streak == 1997 ~ as.character(`1997`),
                     start_fallow_streak == 1998 ~ as.character(`1998`),
                     start_fallow_streak == 1999 ~ as.character(`1999`),
                     start_fallow_streak == 2000 ~ as.character(`2000`),
                     start_fallow_streak == 2001 ~ as.character(`2001`),
                     start_fallow_streak == 2002 ~ as.character(`2002`),
                     start_fallow_streak == 2003 ~ as.character(`2003`),
                     start_fallow_streak == 2004 ~ as.character(`2004`),
                     start_fallow_streak == 2005 ~ as.character(`2005`),
                     start_fallow_streak == 2006 ~ as.character(`2006`),
                     start_fallow_streak == 2007 ~ as.character(`2007`),
                     start_fallow_streak == 2008 ~ as.character(`2008`),
                     start_fallow_streak == 2009 ~ as.character(`2009`),
                     start_fallow_streak == 2010 ~ as.character(`2010`),
                     start_fallow_streak == 2011 ~ as.character(`2011`),
                     start_fallow_streak == 2012 ~ as.character(`2012`),
                     start_fallow_streak == 2013 ~ as.character(`2013`),
                     start_fallow_streak == 2014 ~ as.character(`2014`),
                     start_fallow_streak == 2015 ~ as.character(`2015`),
                     start_fallow_streak == 2016 ~ as.character(`2016`),
                     start_fallow_streak == 2017 ~ as.character(`2017`),
                     start_fallow_streak == 2018 ~ as.character(`2018`),
                     start_fallow_streak == 2019 ~ as.character(`2019`),
                     start_fallow_streak == 2020 ~ as.character(`2020`),
                                    T ~ as.character(`2021`)))

#cleaning up cols
last_crop_clean <- last_crop %>% 
  mutate(last_crop_before_streak = 
           case_when(longest_fallow_streak == "0" ~ "never_fallowed",
                     T ~ as.character(last_crop_before_streak))) %>% 
  mutate(end_fallow_streak = 
           case_when(longest_fallow_streak == "0" ~ NA_character_,
                     T ~ as.character(end_fallow_streak))) %>% 
   mutate(start_fallow_streak = 
           case_when(longest_fallow_streak == "0" ~ NA_character_,
                     T ~ as.character(start_fallow_streak))) %>% 
  select(-id)

#unnest listed pmt_site so they are each in own column
pivot_clean_data <- last_crop_clean %>% 
  unnest_wider(c(`1997`:`2021`), names_sep = "_")

#store as sf
pivot_clean_sf <- pivot_clean_data %>% 
  st_as_sf()

#write to csv
write_csv(pivot_clean_data, "data/shapefiles_written/1997-2021_pivoted_clean.csv")

#write to shapefile
st_write(pivot_clean_data, "data/shapefiles_written/1997-2021_pivoted_clean.shp")


```


