---
title: "Kern County Fields Through Time"
author: "Iris Foxfoot"
date: "1/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r get_packages}
knitr::opts_chunk$set(echo = TRUE)
library(sf) #for polygons
library(tidyverse) #for data wrangling
library(here) #for file paths
library(raster) #for rasters
library(fasterize) #for making rasters
library(tmap) #for interactive maps
library(janitor) #for cleaning names
library(stampr) #for space time analysis
library(sp) #spatial polygons, required for stampr to work
library(smoothr) #drops slivers
```

# Introduction

# Read in field data and wrangle it

In this section we read in the field data and wrangle it such that the end product is a single dataframe containing shapefiles for fields in all years. We also need to add a unique ID column and a year column. 

```{r field_data_wrangling}
#read in kern data
#must have the following file path in working directory: data/kernfields/kernYEAR.shp

#create list of years
kern_field_years <- c(2005:2019)

#loop through read-in process
for (i in kern_field_years) {
  filename <- paste0("kernfields_", i)
  wd <- paste0("data/kernfields/kern", i, ".shp")
  assign(filename, read_sf(wd))
}

#loop through listed sf objects and tidy
fields_cleaned_list <- lapply(mget(sprintf("kernfields_%d", 2005:2019)), 
                 function(x) 
                   {
                   
                   clean_names(x) %>%
                     st_make_valid() %>% 
                     st_transform(crs = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
                     dplyr::select(permit,
                                   permittee,
                                   pmt_site,
                                   comm,
                                   s_status,
                                   p_status,
                                   acres,
                                   dt_act)
                   
                     }
                 )


#rbind inputs
all_fields_binded <- do.call(rbind, fields_cleaned_list) %>% 
  st_make_valid()

#make id column, get year
all_fields <- all_fields_binded%>% 
  st_make_valid() %>% 
  rownames_to_column(var="ID") %>% 
  mutate(year = lubridate::year(dt_act)) %>% 
  dplyr::select(-dt_act)

#make group ID based on geometry. About half of geometry is redundant
all_fields_grouped <- all_fields %>% 
  mutate(group = match(geometry, unique(geometry)))

#aggregate based on group
all_feilds_agg <- all_fields_grouped %>% 
  dplyr::select(group, geometry) %>% 
  group_by(group) %>% 
  summarise(geometry = st_union(geometry))

all_feilds_agg <- all_feilds_agg %>% 
  st_as_sf() %>% 
  st_make_valid()

st_write(all_feilds_agg, here("data", "shapefiles_written", "fields_gr_by_geom.shp"), driver = "ESRI Shapefile")

# #recast as multipolygon
# all_fields <- st_cast(all_fields, "MULTIPOLYGON")

# #convert to spatial
# all_fields_sp <- as(all_fields, "Spatial")
```

```{r}
#crop

#select region (based off of farmer)
kern_subset <- kern_19_clean %>% 
  filter(agent == "McMANUS/WILSON,MICHELE/AARIN")

plot(kern_subset["p_status"])

#crop
kern_19_crop <- st_crop(kern_19_clean, kern_subset)

plot(kern_19_crop["p_status"])

kern_18_crop <- st_crop(kern_18_clean, kern_subset)

kern_county_crop <- st_crop(kern_county, kern_subset)

plot(kern_county_crop["FID"])
```

```{r}
# Convert to tibble for a dplyr full join on the geometry column.
kern1819_intersect <- full_join(as_tibble(kern_19_crop), as_tibble(kern_18_crop), by = "geometry")

kern_1819_sf <- st_as_sf(kern1819_intersect)

tmap_mode("view")

tm_shape(kern_1819_sf) +
  tm_polygons(c("pmt_site.x", "pmt_site.y"), legend.show = FALSE)
```

```{r}
tm_shape(kern_18_crop) +
  tm_polygons("pmt_site", legend.show = FALSE)
```

```{r}
tm_shape(kern_19_crop) +
  tm_polygons("pmt_site", legend.show = FALSE)
```


STAMP interp

ID1 Polygon ID from T1 polygons; NA if it did not exist,
ID2 Polygon ID from T2 polygons; NA if it did not exist,
LEV1 Level 1 STAMP designation,
LEV2 Level 2 STAMP designation,
LEV3 Level 3 STAMP designation,
LEV4 Level 4 STAMP designation,
GROUP Group ID signifying group membership,
AREA Polygon area in appropriate areal units,
-- (optional) Additional columns from directional analysis if direction = TRUE,
-- (optional) Additional columns from distance analysis if distance = TRUE,


STAMP events are reported at four levels of increasing complexity:
LEV1 – disappearance (DISA), stable (STBL), and generation (GENA);
LEV2 – disappearance (DISA), contraction (CONT), stable (STBL), expansion (EXPN), and generation
(GENR);
LEV3 – disappearance (DISA), T1 displacement (DISP1), convergence (CONV), concentration (CONC),
contraction (CONT), stable (STBL), expansion (EXP), fragmentation (FRAG), divergence (DIV), T2 displacement (DISP2), and generation (GENR);
LEV4 – LEV4 is different from other levels. It is used to identify those groups where union (UNION),
division (DIVISION), and both union and division (BOTH) events occur. 




